Synthesizing Quotient Lenses

### 2018-03-08 papers/QuotientLenses/main.tex
Responses to comments and updates to figure -- holding no locks

\sam{Here is my response to this. You observe that the idea of using half a lens
as a canonizer comes out of nowhere, and that this ideas is a hack and always
was. I don't think that's quite right. I think that it is useful to have a
combinator which chooses between two or more formats. Indeed, the permutation
combinator is just a special case of this. The bibtex name field is a good
example of where such a combinator would be useful.  But if we choose to have
such a combinator, then we need a way of mapping one format to another, and
really the only interesting string-to-string transformations in Boomerang are
given by lenses, which is why we use the get function of a lens in a canonizer.
We could have used any other function between the two formats but Boomerang
doesn't give us too many options for writing unidirectional transformations. So
I don't think that using a lens as a canonizer is a trick. I think that
Boomerang just isn't designed for unidirectional tranformations. But I also
agree that it doesn't make much sense to synthesize a lens only to get rid of
one half of it.
I'm not sure what the solution is to be honest.}

@@

Indeed this basic idea is what motivates our definition of {\em QRE Lenses}.
Intuitively, just like Boomerang's quotient lenses, our QRE lenses are bijective
lenses with ``canonizers at the edges''---Figure~\ref{fig:attheedges} presents
the architecture of each lens.
\begin{figure}[t]
\centering
\includegraphics[width=0.7\textwidth]{canonizers-outside}
\caption{Quotient Lens with ``Canonizers at the Edges''
\sam{We don't need a general choose function. With QREs the kernel language is
always contained in the whole language so we just use the inclusion as the
choose function.}\bcp{Done.  But I wonder if the picture should mention QREs
explicitly, somehow... any ideas?}}\sam{Maybe we can put all structures, the
canonizer and the canonical strcutre in a box that says QRE?}
\label{fig:attheedges}
\end{figure}
Every QRE lens $q$ has a type $q: c_1 \Leftrightarrow c_2$ where $c_1$ and $c_2$
are QREs. In the get direction, a QRE lens $q: c_1 \Leftrightarrow c_2$ uses the
source QRE $c_1$ to compute a unique representative for the data modulo the
equivalence relation defined by $c_1$ and then applies the $\get$ function of a
bijective lens $\ell$ to this representative. In the backward direction, $q$
operates similarly, but using the view QRE $c_2$ and the $\lput$ function of
$\ell$.

@@

The formal semantics of QREs satisfy the following theorem:
\begin{theorem}
If $c$ is a well formed QRE, then
\begin{enumerate}
  \item $W(c)$ and $K(c)$ are well defined regular expressions,
  \item  $\eqrel{c}$ is an equivalence relation on $\mathcal{L}(W(c))$,
  \item  $\mathcal{L}(K(c))$ forms a complete set of representatives for
  $\eqrel{c}$,
  \item $\canonizer(c):\mathcal{L}(W(c)) \longrightarrow \mathcal{L}(K(c))$ is
  a well-defined function, and
  \item  given any $w \in \mathcal{L}(W(c))$, $\canonizer(c)(w)$ is the unique
  $k$ in $\mathcal{L}(K(c))$ such that $k$ is equivalent to $w$ modulo
  $\eqrel{c}$.
\end{enumerate}
\end{theorem}

>>>>

\sam{Here is my response to this. You observe that the idea of using half a lens
as a canonizer comes out of nowhere, and that this ideas is a hack and always
was. I don't think that's quite right. I think that it is useful to have a
combinator which chooses between two or more formats. Indeed, the permutation
combinator is just a special case of this. The bibtex name field is a good
example of where such a combinator would be useful.  But if we choose to have
such a combinator, then we need a way of mapping one format to another, and
really the only interesting string-to-string transformations in Boomerang are
given by lenses, which is why we use the get function of a lens in a canonizer.
We could have used any other function between the two formats but Boomerang
doesn't give us too many options for writing unidirectional transformations. So
I don't think that using a lens as a canonizer is a trick. I think that
Boomerang just isn't designed for unidirectional tranformations. But I also
agree that it doesn't make much sense to synthesize a lens only to get rid of
one half of it.
I'm not sure what the solution is to be honest.}
\bcp{Maybe just admit a bit more frankly that this is a convenient trick
  to avoid introducing a whole new language for writing (or synthesizing)
  unidirectional functions.}

@@

Indeed this basic idea is what motivates our definition of {\em QRE Lenses}.
Intuitively, just like Boomerang's quotient lenses, our QRE lenses are bijective
lenses with ``canonizers at the edges''---Figure~\ref{fig:attheedges} presents
the architecture of each lens.
\begin{figure}[t]
\centering
\includegraphics[width=0.7\textwidth]{canonizers-outside}
\caption{Quotient Lens with ``Canonizers at the Edges''
\sam{We don't need a general choose function. With QREs the kernel language is
always contained in the whole language so we just use the inclusion as the
choose function.}\bcp{Done.  But I wonder if the picture should mention QREs
explicitly, somehow... any ideas?}}\sam{Maybe we can put all structures, the
canonizer and the canonical strcutre in a box that says QRE?}\bcp{Like this?
 (But maybe we should go a step further and label things like $K(c)$,
 $W(c)$, Canonizer(c), etc.  And probably we should display K as a subset of
 W.  I can make these changes later.)}
\label{fig:attheedges}
\end{figure}
Every QRE lens $q$ has a type $q: c_1 \Leftrightarrow c_2$ where $c_1$ and $c_2$
are QREs. In the get direction, a QRE lens $q: c_1 \Leftrightarrow c_2$ uses the
source QRE $c_1$ to compute a unique representative for the data modulo the
equivalence relation defined by $c_1$ and then applies the $\get$ function of a
bijective lens $\ell$ to this representative. In the backward direction, $q$
operates similarly, but using the view QRE $c_2$ and the $\lput$ function of
$\ell$.

@@

The formal semantics of QREs satisfy the following theorem:\bcp{That bit is
  entirely redundant with the theorem itself.  Either simply delete it or
  replace with something more interesting. :-)}
\begin{theorem}
If $c$ is a well formed QRE, then
\begin{enumerate}
  \item $W(c)$ and $K(c)$ are well defined regular expressions,
  \item  $\eqrel{c}$ is an equivalence relation on $\mathcal{L}(W(c))$,
  \item  $\mathcal{L}(K(c))$ forms a complete set of representatives for
  $\eqrel{c}$,
  \item $\canonizer(c):\mathcal{L}(W(c)) \longrightarrow \mathcal{L}(K(c))$ is
  a well-defined function, and
  \item  given any $w \in \mathcal{L}(W(c))$, $\canonizer(c)(w)$ is the unique
  $k$ in $\mathcal{L}(K(c))$ such that $k$ is equivalent to $w$ modulo
  $\eqrel{c}$.
\end{enumerate}
\end{theorem}
